\section{VC Dimension}
\label{sec:vc-dimension}
\begin{enumerate}
\item ~[5 points] Assume that the three points below can be labeled
  in any way.  Show with pictures how they can be shattered by a
  linear classifier.  Use filled dots to represent positive classes
  and unfilled dots to represent negative classes.


  \begin{tikzpicture}
    \begin{axis}[my style, xtick={-1,0,...,3}, ytick={-1,0,...,3},
      xmin=-1, xmax=3, ymin=-1, ymax=3]
      \addplot[mark=*,only marks] coordinates {(2,2)(1,1)(1,2)};
    \end{axis}
  \end{tikzpicture}
  
\item {\bf VC-dimension of axis aligned rectangles in $\mathbb{R}^d$}:
  Let $H^d_{rec}$ be the class of axis-aligned rectangles in
  $\mathbb{R}^d$. When $d=2$, this class simply consists of rectangles
  on the plane, and labels all points strictly outside the rectangle
  as negative and all points on or inside the rectangle as positive.
  In higher dimensions, this generalizes to $d$-dimensional boxes,
  with points outside the box labeled negative.

  \begin{enumerate}
  \item ~[10 points] Show that the VC dimension of $H^2_{rec}$ is 4.
  \item ~[10 points] Generalize your argument from the previous proof
    to show that for $d$ dimensions, the VC dimension of $H^d_{rec}$
    is $2d$.
  \end{enumerate}
  
\item In the lectures, we considered the VC dimensions of infinite
  concept classes. However, the same argument can be applied to finite
  concept classes too. In this question, we will explore this setting.

  \begin{enumerate}
  \item ~[10 points] Show that for a finite hypothesis class
    $\mathcal{C}$, its VC dimension can be at most
    $\log_2\left(|\mathcal{C}|\right)$. (Hint: You can use
    contradiction for this proof. But not necessarily!)

  \item ~[5 points] Find an example of a class $\mathcal{C}$ of
    functions over the real interval $X = [0,1]$ such that
    $\mathcal{C}$ is an {\bf infinite} set, while its VC dimension is
    exactly one.

  \item ~[5 points] Give an example of a {\bf finite} class
    $\mathcal{C}$ of functions over the same domain $X = [0,1]$ whose
    VC dimension is exactly $\log_2(|\mathcal{C}|)$.

  \end{enumerate}
  
\end{enumerate}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "hw3"
%%% End:
